---
title: "Data from radiation and weather in Viikki field"
subtitle: "Hourly means, 2024-"
author: "Pedro J. Aphalo"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
# library(tint)
# invalidate cache when the package version changes
knitr::opts_chunk$set(tidy = FALSE) # , cache.extra = packageVersion('tint')
options(htmltools.dir.version = FALSE)
```

When new data is available, first run 'read-data.Rmd', possibly after clearing the 'knitr' cache.

```{r}
library(tidyverse)
library(ggridges)
library(lubridate)

options(warnPartialMatchAttr = FALSE,
        warnPartialMatchDollar = FALSE,
        warnPartialMatchArgs = FALSE)
```


```{r}
rm(list = ls(pattern = "*"))
load("data-rda/hour_2024_latest.tb.rda")
range(hour_2024_latest.tb$time)
colnames(hour_2024_latest.tb)
print(object.size(hour_2024_latest.tb), standard = "SI", units = "MB", digits = 2)
```

## Important note

ggplot stores the whole data frame in the ggplot object. Here we have 203 variables and will in the future possibly have thousands of observations. I have changed the code to first select the columns to be used before building the plot. This reduces the size of plots from 72 MB to 2-3 MB and also makes plotting a lot faster.

## Density plots by month (ridge plots)

## PAR photon irradiance

One hour averages for sun above the horizon of measurements with LI-190 quantum sensor every 5 seconds. During winter-time some snow could have accumulated on the sensor.


```{r}
hour_2024_latest.tb |>
  filter(PAR_umol_mean > 0.01) |>
  select(month_name, PAR_umol_mean, calendar_year) |>
  ggplot(aes(x = PAR_umol_mean, y = month_name)) +
  geom_density_ridges(na.rm = TRUE, scale = 2.5, fill = "darkgreen",
                      alpha = 0.33, bandwidth = 0.05) +
  scale_x_log10(breaks = c(0.01, 0.03, 0.1, 0.3, 1, 3, 10, 30, 100, 300, 1000, 3000), 
                minor_breaks = NULL, 
                labels = scales::format_format(digits = 3, scientific = FALSE)) +
  scale_y_discrete(drop=FALSE) +
#  coord_cartesian(xlim = c(1, 3000)) +
  labs(x = expression(PAR~~photon~~irradiance~~(mu*mol~m^{-2}~s^{-1})), y = "Month of the year") -> par_month_rigdes.fig
par_month_rigdes.fig
```

```{r, eval=FALSE}
par_month_rigdes.fig +
  facet_wrap(~calendar_year) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
```


```{r}
pdf("figures-pdf/par-month-ridges-fig.pdf", width = 10, height = 6)
print(par_month_rigdes.fig)
dev.off()
```

## ET reference

One minute averages for sun above the horizon of measurements with LI-190 quantum sensor every 5 seconds. During winter-time some snow could have accumulated on the sensor.

```{r}
hour_2024_latest.tb |>
  filter(PAR_umol_mean > 5) |>
 select(month_name, ET_ref_FAO56_mean) |>
 ggplot(aes(x = ET_ref_FAO56_mean, y = month_name)) +
 geom_vline(xintercept = c(0.1, 1), linetype = "dashed", colour = "red") +
 geom_density_ridges(na.rm = TRUE, scale = 4.5, fill = "grey50", alpha = 0.33, bandwidth = 0.01) +
 scale_y_discrete(drop=FALSE) +
 scale_x_continuous(breaks = c(0, 0.1, 0.25, 0.5, 0.75, 1)) +
 labs(x = "Evapotranspiration FAO56", y = "Month of the year") -> ET_ref_FAO_month_rigdes.fig
 ET_ref_FAO_month_rigdes.fig
```

```{r}
pdf("figures-pdf/ET-ref-FAO-month-ridges-fig.pdf", width = 10, height = 6)
print(ET_ref_FAO_month_rigdes.fig)
dev.off()
```

```{r}
hour_2024_latest.tb |>
#  filter(sun_elevation > 0 & PAR_Den_Avg > 5) |>
  select(month_name, air_DP_mean) |>
  ggplot(aes(x = air_DP_mean, y = month_name)) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "red") +
  geom_density_ridges(na.rm = TRUE, scale = 3.5, fill = "grey50", alpha = 0.33) +
  scale_y_discrete(drop=FALSE) +
#  scale_x_continuous(breaks = c(0, 0.1, 0.25, 0.5, 0.75, 1)) +
  labs(x = "Dew point (C)", y = "Month of the year") -> dew_point_month_rigdes.fig
dew_point_month_rigdes.fig
```

```{r}
hour_2024_latest.tb |>
#  filter(sun_elevation > 0 & PAR_Den_Avg > 5) |>
  select(month_name, air_temp_C_mean) |>
  ggplot(aes(x = air_temp_C_mean, y = month_name)) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "red") +
  geom_density_ridges(na.rm = TRUE, scale = 3.5, fill = "grey50", alpha = 0.33) +
  scale_y_discrete(drop=FALSE) +
#  scale_x_continuous(breaks = c(0, 0.1, 0.25, 0.5, 0.75, 1)) +
  labs(x = "Air temperature (C)", y = "Month of the year") -> air_temp_month_rigdes.fig
air_temp_month_rigdes.fig
```

```{r}
hour_2024_latest.tb |>
#  filter(sun_elevation > 0 & PAR_Den_Avg > 5) |>
  select(month_name, air_RH_mean) |>
  ggplot(aes(x = air_RH_mean, y = month_name)) +
  geom_vline(xintercept = c(0, 100), linetype = "dashed", colour = "red") +
  geom_density_ridges(na.rm = TRUE, scale = 3.5, fill = "grey50", alpha = 0.33) +
  scale_y_discrete(drop=FALSE) +
#  scale_x_continuous(breaks = c(0, 0.1, 0.25, 0.5, 0.75, 1)) +
  labs(x = "Air relative humidity (%)", y = "Month of the year") -> air_temp_month_rigdes.fig
air_temp_month_rigdes.fig
```

```{r}
hour_2024_latest.tb |>
#  filter(sun_elevation > 0 & PAR_Den_Avg > 5) |>
  select(month_name, wind_speed_mean) |>
  ggplot(aes(x = wind_speed_mean, y = month_name)) +
  geom_density_ridges(na.rm = TRUE, scale = 1.5, fill = "grey50", alpha = 0.33) +
  scale_y_discrete(drop=FALSE) +
#  scale_x_continuous(breaks = c(0, 0.1, 0.25, 0.5, 0.75, 1)) +
  labs(x = "Wind speed (m s-1)", y = "Month of the year") -> wind_speed_month_rigdes.fig
wind_speed_month_rigdes.fig
```

```{r}
hour_2024_latest.tb |>
#  filter(sun_elevation > 0 & PAR_Den_Avg > 5) |>
  select(month_name, air_pressure_mean) |>
  ggplot(aes(x = air_pressure_mean, y = month_name)) +
  geom_density_ridges(na.rm = TRUE, scale = 1.5, fill = "grey50", alpha = 0.33) +
  scale_y_discrete(drop=FALSE) +
#  scale_x_continuous(breaks = c(0, 0.1, 0.25, 0.5, 0.75, 1)) +
  labs(x = "Air Pressure (Pa)", y = "Month of the year") -> 
air_pressure_month_rigdes.fig
air_pressure_month_rigdes.fig
```

# Time-series plots

```{r}
hour_2024_latest.tb |>
#  filter(sun_elevation > 0 & PAR_Den_Avg > 5) |>
  select(time, global_watt_mean, global_watt_min, global_watt_max) |>
  ggplot(aes(time)) +
  geom_line(aes(y = global_watt_mean)) +
  geom_line(aes(y = global_watt_min), colour = "blue") +
  geom_line(aes(y = global_watt_max), colour = "red") +
  labs(y = expression("Global radiation "*(W~m^{-2})), x = "Time") -> 
global_radiation.fig
global_radiation.fig
```


```{r}
hour_2024_latest.tb |>
#  filter(sun_elevation > 0 & PAR_Den_Avg > 5) |>
  select(time, PAR_umol_mean, PAR_umol_min, PAR_umol_max) |>
  na.omit() |>
  ggplot(aes(time)) +
  geom_line(aes(y = PAR_umol_mean)) +
  geom_line(aes(y = PAR_umol_min), colour = "blue") +
  geom_line(aes(y = PAR_umol_max), colour = "red") +
  labs(y = expression("PAR "*(mu*mol~m^{-2}~s^{-1})), x = "Time") -> 
PAR.fig
PAR.fig
```

```{r}
hour_2024_latest.tb %>%
#  filter(sun_elevation > 0 & PAR_Den_Avg > 5) %>%
  select(time, UVB_umol_mean, UVB_umol_min, UVB_umol_max) %>%
  ggplot(aes(time)) +
  geom_line(aes(y = UVB_umol_mean)) +
  geom_line(aes(y = UVB_umol_min), colour = "blue") +
  geom_line(aes(y = UVB_umol_max), colour = "red") +
  labs(y = expression("UV-B "*(mu*mol~m^{-2}~s^{-1})), x = "Time") -> 
UVB.fig
UVB.fig
```

```{r}
hour_2024_latest.tb |>
#  filter(sun_elevation > 0 & PAR_Den_Avg > 5) |>
  select(time, UVA2_umol_mean) |>
  ggplot(aes(time, UVA2_umol_mean)) +
  geom_line() +
  labs(y = expression("UV-A2 "*(mu*mol~m^{-2}~s^{-1})), x = "Time") -> 
UVA2.fig
UVA2.fig
```

```{r}
hour_2024_latest.tb |>
#  filter(sun_elevation > 0 & PAR_Den_Avg > 5) |>
  select(time, UVA1_umol_mean) |>
  ggplot(aes(time, UVA1_umol_mean)) +
  geom_line() +
  labs(y = expression("UV-A1 "*(mu*mol~m^{-2}~s^{-1})), x = "Time") -> 
UVA1.fig
UVA1.fig
```

```{r}
hour_2024_latest.tb |>
#  filter(sun_elevation > 0 & PAR_Den_Avg > 5) |>
  select(time, blue_umol_mean) |>
  ggplot(aes(time, blue_umol_mean)) +
  geom_line() +
  labs(y = expression("Blue light "*(mu*mol~m^{-2}~s^{-1})), x = "Time") -> 
Blue.fig
Blue.fig
```

```{r}
hour_2024_latest.tb |>
#  filter(sun_elevation > 0 & PAR_Den_Avg > 5) |>
  select(time, rain_mm_h) |>
  ggplot(aes(time, rain_mm_h)) +
  geom_line() +
  labs(y = expression("Rain "*(mm~h^{-1})), x = "Time") -> 
Blue.fig
Blue.fig
```

```{r}
hour_2024_latest.tb %>%
#  filter(sun_elevation > 0 & PAR_Den_Avg > 5) %>%
  select(time, ET_ref_FAO56_mean, ET_ref_FAO56_min, ET_ref_FAO56_max) %>%
  ggplot(aes(time)) +
  geom_line(aes(y = ET_ref_FAO56_min), colour = "blue") +
  geom_line(aes(y = ET_ref_FAO56_max), colour = "red") +
  geom_line(aes(y = ET_ref_FAO56_mean)) +
  labs(y = expression("Reference ET, FAO56 "*(mm~h^{-1})), x = "Time") -> 
ET_ref_FAO56.fig
ET_ref_FAO56.fig
```

```{r}
hour_2024_latest.tb %>%
#  filter(sun_elevation > 0 & PAR_Den_Avg > 5) %>%
  select(time, R_rel_mean, R_rel_min, R_rel_max) %>%
  ggplot(aes(time)) +
  geom_line(aes(y = R_rel_min), colour = "blue") +
  geom_line(aes(y = R_rel_max), colour = "red") +
  geom_line(aes(y = R_rel_mean)) +
  labs(y = "Relative atmosphere transmittance (/1)", x = "Time") -> 
R_rel.fig
R_rel.fig
```


```{r}
hour_2024_latest.tb %>%
#  filter(sun_elevation > 0 & PAR_Den_Avg > 5) %>%
  select(time, R_0_mean, R_0_min, R_0_max) %>%
  ggplot(aes(time)) +
  geom_line(aes(y = R_0_min), colour = "blue") +
  geom_line(aes(y = R_0_max), colour = "red") +
  geom_line(aes(y = R_0_mean)) +
  labs(y = expression("AM0 radiation (estimate) "*(W~m^{-2})), x = "Time") -> 
R_rel.fig
R_rel.fig
```

```{r}
hour_2024_latest.tb %>%
#  filter(sun_elevation > 0 & PAR_Den_Avg > 5) %>%
  select(time, ET_ref_FAO56_mean, ET_ref_FAO56_min, ET_ref_FAO56_max) %>%
  ggplot(aes(time)) +
  geom_line(aes(y = ET_ref_FAO56_min), colour = "blue") +
  geom_line(aes(y = ET_ref_FAO56_max), colour = "red") +
  geom_line(aes(y = ET_ref_FAO56_mean)) +
  labs(y = expression("Reference ET, FAO56 "*(mm~h^{-1})), x = "Time") -> 
ET_ref_FAO56.fig
ET_ref_FAO56.fig
```
```{r}
hour_2024_latest.tb |>
#  filter(sun_elevation > 0 & PAR_Den_Avg > 5) |>
  select(time, T_5cm, T_10cm, T_20cm, T_30cm, T_40cm, T_50cm) |>
  ggplot() +
  geom_line(aes(y = T_5cm, x = time), color = "red") +
  geom_line(aes(y = T_10cm, x = time)) +
  geom_line(aes(y = T_20cm, x = time), color = "darkgreen") +
  geom_line(aes(y = T_30cm, x = time), color = "darkblue") +
  geom_line(aes(y = T_50cm, x = time), color = "darkred") +
#  scale_x_continuous(breaks = c(0, 0.1, 0.25, 0.5, 0.75, 1)) +
  labs(y = "Soil temperature at 5 to 50 cm depth (C)", x = "Time") -> 
soil_t_50cm.fig
soil_t_50cm.fig
```

```{r}
hour_2024_latest.tb |>
#  filter(sun_elevation > 0 & PAR_Den_Avg > 5) |>
  select(time, VWC_5cm, VWC_10cm, VWC_20cm, VWC_30cm, VWC_50cm) |>
  ggplot() +
  geom_line(aes(y = VWC_5cm, x = time), color = "red") +
  geom_line(aes(y = VWC_10cm, x = time)) +
  geom_line(aes(y = VWC_20cm, x = time), color = "darkgreen") +
  geom_line(aes(y = VWC_30cm, x = time), color = "darkblue") +
  geom_line(aes(y = VWC_50cm, x = time), color = "darkred") +
#  scale_x_continuous(breaks = c(0, 0.1, 0.25, 0.5, 0.75, 1)) +
  labs(y = "Soil water v/v ()", x = "Time") +
  expand_limits(y = 0) -> 
soil_VWC.fig
soil_VWC.fig
```

```{r}
hour_2024_latest.tb |>
#  filter(sun_elevation > 0 & PAR_Den_Avg > 5) |>
  select(time, BulkEC_10cm, BulkEC_20cm, BulkEC_30cm, BulkEC_50cm) |>
  ggplot() +
  geom_line(aes(y = BulkEC_10cm, x = time)) +
  geom_line(aes(y = BulkEC_20cm, x = time), color = "darkgreen") +
  geom_line(aes(y = BulkEC_30cm, x = time), color = "darkblue") +
  geom_line(aes(y = BulkEC_50cm, x = time), color = "darkred") +
#  scale_x_continuous(breaks = c(0, 0.1, 0.25, 0.5, 0.75, 1)) +
  labs(y = "Bulk EC (dS/m)", x = "Time") ->
soil_BulkEC.fig
soil_BulkEC.fig
```

```{r}
hour_2024_latest.tb |>
#  filter(sun_elevation > 0 & PAR_Den_Avg > 5) |>
  select(time, Ka_5cm, Ka_10cm, Ka_20cm, Ka_30cm, Ka_50cm) |>
  ggplot() +
  geom_line(aes(y = Ka_5cm, x = time), color = "red") +
  geom_line(aes(y = Ka_10cm, x = time)) +
  geom_line(aes(y = Ka_20cm, x = time), color = "darkgreen") +
  geom_line(aes(y = Ka_30cm, x = time), color = "darkblue") +
  geom_line(aes(y = Ka_50cm, x = time), color = "darkred") +
#  scale_x_continuous(breaks = c(0, 0.1, 0.25, 0.5, 0.75, 1)) +
  labs(y = "Relative Dielectric Permittivity (1..80)", x = "Time") +
  expand_limits(y = 0) -> 
soil_Ka.fig
soil_Ka.fig
```


## Small plot

```{r}
hour_2024_latest.tb |>
#  filter(sun_elevation > 0 & PAR_Den_Avg > 5) |>
  filter(time >= ymd_hms("2024-06-12 00:00:00")) |>
  select(time, T_1, T_2, T_3) |>
  ggplot() +
  geom_line(aes(y = T_1, x = time), color = "red") +
  geom_line(aes(y = T_2, x = time)) +
  geom_line(aes(y = T_3, x = time), color = "darkblue") +
#  scale_x_continuous(breaks = c(0, 0.1, 0.25, 0.5, 0.75, 1)) +
  labs(y = "Soil temperature at 5 to 50 cm depth (C)", x = "Time") -> 
soil_t_20cm.fig
soil_t_20cm.fig
```


```{r}
hour_2024_latest.tb |>
#  filter(sun_elevation > 0 & PAR_Den_Avg > 5) |>
  filter(time >= ymd_hms("2024-06-12 00:00:00")) |>
  select(time, VWC_1, VWC_2, VWC_3) |>
  ggplot() +
  geom_line(aes(y = VWC_1, x = time), color = "red") +
  geom_line(aes(y = VWC_2, x = time)) +
  geom_line(aes(y = VWC_3, x = time), color = "darkblue") +
#  scale_x_continuous(breaks = c(0, 0.1, 0.25, 0.5, 0.75, 1)) +
  labs(y = "Soil water v/v ()", x = "Time") +
  expand_limits(y = 0) ->
soil_VWC_20cm.fig
soil_VWC_20cm.fig
```


```{r}
hour_2024_latest.tb |>
  filter(time >= ymd_hms("2024-06-12 00:00:00")) |>
  select(time, EC_1, EC_2, EC_3) |>
  ggplot() +
  geom_line(aes(y = EC_1, x = time), color = "red") +
  geom_line(aes(y = EC_2, x = time)) +
  geom_line(aes(y = EC_3, x = time), color = "darkblue") +
    labs(y = "Bulk EC (dS/m)", x = "Time") -> 
soil_t_50cm.fig
soil_t_50cm.fig
```
