---
title: "Read data from Viikki field weather station"
subtitle: "50-ms-interval logging with Campbell Scientific CR6"
author: "Pedro J. Aphalo"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(tidy = FALSE, echo = TRUE)
```

# News

Logging started to 50 ms on 2024-04-15.

For 50 ms to work the notch frequency had to be increased from to 500 Hz.
This decreased the noise-reducing averaging per individual data 
acquisition+conversion from approximately 200 ms to 20 ms.

# Import and preprocessing of the data logged at 1 s intervals

```{r}
library(readr)
library(lubridate)
library(photobiology)
library(photobiologyInOut)
library(ggspectra)
library(lubridate)
library(dplyr)
library(ggpp)
```

We store the coordinates of the station to within < 5 m, plus any possible error in Google Maps image layer coordinates.

```{r}
viikki_bio3.geo <- data.frame(lon = 25.019212, lat = 60.226805, address = "BIO3, Viikki")
locale_UTC <- locale(tz = "UTC")
```

We read the 50-ms-interval data for whole period. These data are only for and for 5 min for each hour, from 5 to 20 min past the hour. 

When downloading data from the logger we append it to the existing file. A new file is started when the data tables change, i.e., when the edits to the logger program modifies the structure of a data table. This resetting is done per table. From time to time, such as once per year, I force a flushing of old data, even when data tables have not changed. 

```{r}
millisecond_raw.tb <- read_csi_dat(file = "data-latest/Viikki Tower_TableMilliSecond.dat", locale = locale_UTC)

millisecond_raw.tb
nrow(millisecond_raw.tb)
ncol(millisecond_raw.tb)
```

```{r}
colnames(millisecond_raw.tb)
```

```{r}
cat(comment(millisecond_raw.tb))
```

Sometimes after changes in the logger program data already downloaded has been
appended to files, leading to duplicate rows.

```{r}
rle.check <- rle(sort(as.numeric(millisecond_raw.tb$TIMESTAMP)))
duplicates <- sum(rle.check$lengths > 1)
if (duplicates > 0L) {
  message("Found ", duplicates, " duplicated rows in data.")
  millisecond_raw.tb <- distinct(millisecond_raw.tb, TIMESTAMP, .keep_all = TRUE)
}
```

To force TZ with 'lubridate' we would need to be sure that daylight saving times have been automatically set. This is not the case so we need first to subtract 2 h from UTC + 2h time, to convert to true UTC and then do calculations based on this. We re-express all times in UTC year round.

```{r}
millisecond_raw.tb[["TIMESTAMP"]][1]
millisecond_raw.tb[["TIMESTAMP"]][nrow(millisecond_raw.tb)]
tz(millisecond_raw.tb[["TIMESTAMP"]][1])
millisecond_raw.tb[["TIMESTAMP"]] <-  millisecond_raw.tb[["TIMESTAMP"]] - hours(2) # UTC + 2h -> UTC
millisecond_raw.tb[["TIMESTAMP"]][1]
millisecond_raw.tb[["TIMESTAMP"]][nrow(millisecond_raw.tb)]
tz(millisecond_raw.tb[["TIMESTAMP"]][1])
```

In the next code chunk we do the following:

a. Extract components of date time.
b. Compute solar time and sun elevation and azimuth.
FALSE) or not (sunny = TRUE).

```{r}
  # NEW CR6 program
millisecond_raw.tb %>%
  filter(!is.na(TIMESTAMP)) %>%
  mutate(year = year(TIMESTAMP),
         month_of_year = month(TIMESTAMP),
         month_name = factor(month_of_year, levels = 12:1, labels = rev(month.name)),
         week_of_year = week(TIMESTAMP),
         day_of_year = yday(TIMESTAMP),
         time_of_day_utc = as_tod(TIMESTAMP),
         solar_time_h = solar_time(TIMESTAMP, 
                                   geocode = viikki_bio3.geo),
         solar_time_s = solar_time(TIMESTAMP, 
                                   geocode = viikki_bio3.geo, 
                                   unit.out = "seconds"),
         sun_elevation = sun_elevation(TIMESTAMP, 
                                       geocode = viikki_bio3.geo, 
                                       use.refraction = TRUE,
                                       tz = "UTC"),
         sun_azimuth = sun_azimuth(TIMESTAMP, 
                                   geocode = viikki_bio3.geo, 
                                   use.refraction = TRUE,
                                   tz = "UTC")) -> millisecond.tb
```

```{r}
colnames(millisecond.tb)
```

```{r}
obj.name <- paste("millisecond_", year(millisecond_raw.tb$TIMESTAMP[1]),
                  "_", month(millisecond_raw.tb$TIMESTAMP[1]), 
                  "_", day(millisecond_raw.tb$TIMESTAMP[1]), ".tb", sep = "")
assign(obj.name, millisecond.tb)

save(list = obj.name, file = paste("data-rda/", obj.name, ".rda", sep = ""))
```


```{r}
millisecond.tb %>%
  head()
millisecond.tb %>%
  tail()
```

```{r}
ggplot(millisecond.tb[-1, ], aes(PAR_Den_CS)) +
  stat_density() +
  scale_x_continuous(limits = function(x) {c(x[1] - 100, x[2] + 20)})
```

```{r}
ggplot(millisecond.tb[-1, ], aes(TIMESTAMP, PAR_Den_CS)) +
  geom_line() +
  expand_limits(y = 0)
```


```{r}
ggplot(subset(millisecond.tb[-1, ], TIMESTAMP > ymd_hm("2024-06-07 1:00") & TIMESTAMP < ymd_hm("2024-06-07 1:15")), aes(TIMESTAMP, PAR_Den_CS)) +
  geom_line() +
  stat_panel_counts(label.y = "bottom") +
  expand_limits(y = 0)

```

```{r}
ggplot(subset(millisecond.tb[-1, ], TIMESTAMP > ymd_hm("2024-06-07 2:00") & TIMESTAMP < ymd_hm("2024-06-07 2:15")), aes(TIMESTAMP, PAR_Den_CS)) +
  geom_line() +
  stat_panel_counts(label.y = "bottom") +
  expand_limits(y = 0)

```

```{r}
ggplot(subset(millisecond.tb[-1, ], TIMESTAMP > ymd_hm("2024-06-07 3:00") & TIMESTAMP < ymd_hm("2024-06-07 3:15")), aes(TIMESTAMP, PAR_Den_CS)) +
  geom_line() +
  stat_panel_counts(label.y = "bottom") +
  expand_limits(y = 0)

```

```{r}
ggplot(subset(millisecond.tb[-1, ], TIMESTAMP > ymd_hm("2024-06-07 4:00") & TIMESTAMP < ymd_hm("2024-06-07 4:15")), aes(TIMESTAMP, PAR_Den_CS)) +
  geom_line() +
  stat_panel_counts(label.y = "bottom") +
  expand_limits(y = 0)

```


```{r}
ggplot(subset(millisecond.tb[-1, ], TIMESTAMP > ymd_hm("2024-06-07 5:00") & TIMESTAMP < ymd_hm("2024-06-07 5:15")), aes(TIMESTAMP, PAR_Den_CS)) +
  geom_line() +
  stat_panel_counts(label.y = "bottom") +
  expand_limits(y = 0)

```

```{r}
ggplot(subset(millisecond.tb[-1, ], TIMESTAMP > ymd_hm("2024-06-07 6:00") & TIMESTAMP < ymd_hm("2024-06-07 6:15")), aes(TIMESTAMP, PAR_Den_CS)) +
  geom_line() +
  stat_panel_counts(label.y = "bottom") +
  expand_limits(y = 0)

```


```{r}
ggplot(subset(millisecond.tb[-1, ], TIMESTAMP > ymd_hm("2024-06-07 7:00") & TIMESTAMP < ymd_hm("2024-06-07 7:15")), aes(TIMESTAMP, PAR_Den_CS)) +
  geom_line() +
  stat_panel_counts(label.y = "bottom") +
  expand_limits(y = 0)
```
`

```{r}
ggplot(subset(millisecond.tb[-1, ], TIMESTAMP > ymd_hm("2024-06-07 8:00") & TIMESTAMP < ymd_hm("2024-06-07 8:15")), aes(TIMESTAMP, PAR_Den_CS)) +
  geom_line() +
  stat_panel_counts(label.y = "bottom") +
  expand_limits(y = 0)

```

```{r}
ggplot(subset(millisecond.tb[-1, ], TIMESTAMP > ymd_hm("2024-06-07 9:00") & TIMESTAMP < ymd_hm("2024-06-07 9:15")), aes(TIMESTAMP, PAR_Den_CS)) +
  geom_line() +
  stat_panel_counts(label.y = "bottom") +
  expand_limits(y = 0)

```

```{r}
ggplot(subset(millisecond.tb[-1, ], TIMESTAMP > ymd_hm("2024-06-07 10:00") & TIMESTAMP < ymd_hm("2024-06-07 10:15")), aes(TIMESTAMP, PAR_Den_CS)) +
  geom_line() +
  stat_panel_counts(label.y = "bottom") +
  expand_limits(y = 0)

```

```{r}
ggplot(subset(millisecond.tb[-1, ], TIMESTAMP > ymd_hm("2024-06-07 11:00") & TIMESTAMP < ymd_hm("2024-06-07 11:15")), aes(TIMESTAMP, PAR_Den_CS)) +
  geom_line() +
  stat_panel_counts(label.y = "bottom") +
  expand_limits(y = 0)

```

```{r}
ggplot(subset(millisecond.tb[-1, ], TIMESTAMP > ymd_hm("2024-06-07 12:00") & TIMESTAMP < ymd_hm("2024-06-07 12:15")), aes(TIMESTAMP, PAR_Den_CS)) +
  geom_line() +
  stat_panel_counts(label.y = "bottom") +
  expand_limits(y = 0)

```

```{r}
ggplot(subset(millisecond.tb[-1, ], TIMESTAMP > ymd_hm("2024-06-07 13:00") & TIMESTAMP < ymd_hm("2024-06-07 13:15")), aes(TIMESTAMP, PAR_Den_CS)) +
  geom_line() +
  stat_panel_counts(label.y = "bottom") +
  expand_limits(y = 0)

```

```{r}
ggplot(subset(millisecond.tb[-1, ], TIMESTAMP > ymd_hm("2024-06-07 14:00") & TIMESTAMP < ymd_hm("2024-06-07 14:15")), aes(TIMESTAMP, PAR_Den_CS)) +
  geom_line() +
  stat_panel_counts(label.y = "bottom") +
  expand_limits(y = 0)

```

```{r}
ggplot(subset(millisecond.tb[-1, ], TIMESTAMP > ymd_hm("2024-06-07 15:00") & TIMESTAMP < ymd_hm("2024-06-07 15:15")), aes(TIMESTAMP, PAR_Den_CS)) +
  geom_line() +
  stat_panel_counts(label.y = "bottom") +
  expand_limits(y = 0)

```

```{r}
ggplot(subset(millisecond.tb[-1, ], TIMESTAMP > ymd_hm("2024-06-07 16:00") & TIMESTAMP < ymd_hm("2024-06-07 16:15")), aes(TIMESTAMP, PAR_Den_CS)) +
  geom_line() +
  stat_panel_counts(label.y = "bottom") +
  expand_limits(y = 0)

```

```{r}
ggplot(subset(millisecond.tb[-1, ], TIMESTAMP > ymd_hm("2024-06-07 17:00") & TIMESTAMP < ymd_hm("2024-06-07 17:15")), aes(TIMESTAMP, PAR_Den_CS)) +
  geom_line() +
  expand_limits(y = 0)

```

```{r}
ggplot(subset(millisecond.tb[-1, ], TIMESTAMP > ymd_hm("2024-06-07 19:00") & TIMESTAMP < ymd_hm("2024-06-07 19:15")), aes(TIMESTAMP, PAR_Den_CS)) +
  geom_line() +
  expand_limits(y = 0)

```


```{r}
ggplot(subset(millisecond.tb[-1, ], TIMESTAMP > ymd_hm("2024-06-07 18:00") & TIMESTAMP < ymd_hm("2024-06-07 18:15")), aes(TIMESTAMP, PAR_Den_CS)) +
  geom_line() +
  stat_panel_counts(label.y = "bottom") +
  expand_limits(y = 0)

```

```{r}
ggplot(subset(millisecond.tb[-1, ], TIMESTAMP > ymd_hms("2024-05-10 15:03:30") & TIMESTAMP < ymd_hms("2024-05-10 15:03:45")), aes(TIMESTAMP, PAR_Den_CS)) +
  geom_line() +
  stat_panel_counts(label.y = "bottom") +
  expand_limits(y = 0)
```
