---
title: "Check 1 s data"
author: "Pedro J. Aphalo"
date: "`r Sys.Date()`"
output: html_document
---

# Read the merged data logged at 1 s intervals

```{r}
library(readr)
library(lubridate)
library(photobiology)
library(photobiologyInOut)
library(ggspectra)
library(lubridate)
library(dplyr)
library(ggplot2)
library(ggpmisc)
library(ggdensity)

theme_set(theme_bw())
```

# Density ditributions

Data at 1 s is logged only if `PAR_umol_CS` > 1 umol m-2 s-1.

```{r}
load("data-rda/second_2024_latest.tb.rda")
ncol(second_2024_latest.tb)
nrow(second_2024_latest.tb)
print(object.size(second_2024_latest.tb), unit = "MB", standard = "SI")
```

```{r}
ggplot(second_2024_latest.tb, aes(PAR_umol)) +
  stat_density() +
  scale_x_continuous(limits = function(x) {c(x[1] - 100, x[2] + 20)})
```

```{r}
ggplot(second_2024_latest.tb, aes(PAR_diff_fr)) +
  stat_density() +
  scale_x_continuous(limits = function(x) {c(x[1] - 0.15, x[2] + 0.15)})
```

```{r}
ggplot(second_2024_latest.tb, aes(PAR_umol_BF)) +
  stat_density() +
  scale_x_continuous(limits = function(x) {c(x[1] - 100, x[2] + 0.20)})
```

# Compare Delta-T BF5 against Apogee/Campbel CS310

Fresh calibration needed for 2024!

```{r}
ggplot(second_2024_latest.tb[ , c("PAR_umol_CS", "PAR_umol_BF")], 
       aes(PAR_umol_CS, PAR_umol_BF)) +
  stat_hdr() +
  stat_ma_line() +
  stat_ma_eq(use_label(c("eq", "R2", "n"))) +
  expand_limits(x = -150, y = -150)
```

Check switch-off of logging at 1 s based on `PAR_umol_CS` < 1 umol m-2 s-1.

```{r}
ggplot(subset(second_2024_latest.tb, 
              time >= ymd_hm("2024-06-04 17:30") & time <= ymd_hm("2024-06-05 03:00")),
       aes(time, PAR_umol_CS)) +
  geom_line(alpha = 0.5) +
  expand_limits(y = 0)
```

BF5 readings are a lot more variable.

```{r}
ggplot(subset(second_2024_latest.tb, 
              time >= ymd_hm("2024-06-04 17:30") & time <= ymd_hm("2024-06-05 03:00")),
       aes(time, PAR_umol_BF)) +
  geom_line(alpha = 0.5)

```

Even more clearly seen here. Thus, BF readings are not much useful at 1 s frequency, except possibly to detect occlusion of the CS sensor by birds, in addition the relationship at low irradiance is slightly curved. Of course, sunshine is probably usable as `sunny`.

```{r}
p1 <-
  ggplot(subset(second_2024_latest.tb, 
              time >= ymd_hm("2024-06-04 13:30") & time <= ymd_hm("2024-06-05 07:00")),
       aes(PAR_umol_CS, PAR_umol_BF, colour = lubridate::hour(time) < 22 & lubridate::hour(time) > 14)) +
  geom_point(alpha = 0.5, size = 0.5) +
  geom_abline(slope = 1, intercept = 0, colour = "blue")
p1
p1 + coord_cartesian(xlim = c(0, 400), ylim = c(0, 400))
```

# Overview

```{r}
ggplot(second_2024_latest.tb, aes(time, PAR_umol)) +
  geom_line()
```

```{r}
ggplot(second_2024_latest.tb, aes(time, far_red_umol)) +
  geom_line()
```

```{r}
ggplot(second_2024_latest.tb, aes(time, UVA_umol)) +
  geom_line()
```

```{r}
ggplot(second_2024_latest.tb, aes(time, UVB_umol)) +
  geom_line()
```

# Diffuse fraction

The measuwhite diffuse fraction is very noisy at the start and end of the day. Making the value of factor `sun_disk` unreliable at very low `sun_elevation` values.

```{r}
ggplot(second_2024_latest.tb, aes(time, PAR_diff_fr)) +
  geom_line() +
  expand_limits(y = 0) +
  scale_y_continuous(breaks = c(0, 0.25, 0.5, 0.75, 1.0))
```

Because of the noisy data from the BF5 and the relatively small difference in the expected diffuse fraction with and without oclusions of the solar disk, we set factor `solar_disk` to `NA` whem `PAR_umol_BF` < 10 umol m-2 s-1. Looking at a single day we can see that partial occlusion of the solar disk, still shown as `"visible"` can markedly decrease irradiance. The missing estimates of solar disk occlusion are shown as white.

```{r}
ggplot(subset(second_2024_latest.tb, 
              time >= ymd_hm("2024-05-05 00:00") & time <= ymd_hm("2024-05-05 23:59")),
       aes(time, PAR_umol)) +
  geom_point(aes(colour = solar_disk), alpha = 0.5) +
  geom_line(alpha = 0.5) +
  scale_colour_manual(breaks = c("occluded", "visible"), 
                      na.value = "white",
                      values = c("grey75", "orange")) +
  expand_limits(y = c(0, 2100))

```

```{r}
ggplot(subset(second_2024_latest.tb, 
              time >= ymd_hm("2024-05-05 00:00") & time <= ymd_hm("2024-05-05 23:59")),
       aes(time, UVB_umol)) +
  geom_point(aes(colour = solar_disk), alpha = 0.5) +
  geom_line(alpha = 0.5) +
  scale_colour_manual(breaks = c("occluded", "visible"), 
                      na.value = "white",
                      values = c("grey75", "orange")) +
  expand_limits(y = c(0, 2.5))

```


```{r}
ggplot(subset(second_2024_latest.tb, 
              time >= ymd_hm("2024-05-05 00:00") & time <= ymd_hm("2024-05-05 23:59")),
       aes(time, far_red_umol)) +
  geom_point(aes(colour = solar_disk), alpha = 0.5) +
  geom_line(alpha = 0.5) +
  scale_colour_manual(breaks = c("occluded", "visible"), 
                      na.value = "white",
                      values = c("grey75", "orange")) +
  expand_limits(y = c(0, 70))

```

```{r}
ggplot(subset(second_2024_latest.tb, 
              time >= ymd_hm("2024-04-30 00:00") & time <= ymd_hm("2024-04-30 23:59")),
       aes(time, PAR_umol)) +
  geom_point(aes(colour = solar_disk), alpha = 0.05) +
  geom_line(alpha = 0.5) +
  scale_colour_manual(breaks = c("occluded", "visible"), 
                      na.value = "white",
                      values = c("grey75", "orange")) +
  expand_limits(y = c(0, 2100))

```


```{r}
ggplot(subset(second_2024_latest.tb, 
              time >= ymd_hm("2024-04-30 00:00") & time <= ymd_hm("2024-04-30 23:59")),
       aes(time, UVB_umol)) +
  geom_point(aes(colour = solar_disk), alpha = 0.5) +
  geom_line(alpha = 0.5) +
  scale_colour_manual(breaks = c("occluded", "visible"), 
                      na.value = "white",
                      values = c("grey75", "orange")) +
  expand_limits(y = c(0, 2.5))

```


```{r}
ggplot(subset(second_2024_latest.tb, 
              time >= ymd_hm("2024-04-30 00:00") & time <= ymd_hm("2024-04-30 23:59")),
       aes(time, far_red_umol)) +
  geom_point(aes(colour = solar_disk), alpha = 0.5) +
  geom_line(alpha = 0.5) +
  scale_colour_manual(breaks = c("occluded", "visible"), 
                      na.value = "white",
                      values = c("grey75", "orange")) +
  expand_limits(y = c(0, 70))

```

```{r}
ggplot(subset(second_2024_latest.tb, 
              time >= ymd_hm("2024-04-22 00:00") & time <= ymd_hm("2024-04-22 23:59")),
       aes(time, PAR_umol)) +
  geom_point(aes(colour = solar_disk), alpha = 0.5) +
  geom_line(alpha = 0.5) +
  scale_colour_manual(breaks = c("occluded", "visible"), 
                      na.value = "white",
                      values = c("grey75", "orange")) +
  expand_limits(y = c(0, 2100))

```


```{r}
ggplot(subset(second_2024_latest.tb, 
              time >= ymd_hm("2024-04-22 00:00") & time <= ymd_hm("2024-04-22 23:59")),
       aes(time, UVB_umol)) +
  geom_point(aes(colour = solar_disk), alpha = 0.5) +
  geom_line(alpha = 0.5) +
  scale_colour_manual(breaks = c("occluded", "visible"), 
                      na.value = "white",
                      values = c("grey75", "orange")) +
  expand_limits(y = c(0, 2.5))

```


```{r}
ggplot(subset(second_2024_latest.tb, 
              time >= ymd_hm("2024-04-22 00:00") & time <= ymd_hm("2024-04-22 23:59")),
       aes(time, far_red_umol)) +
  geom_point(aes(colour = solar_disk), alpha = 0.05) +
  geom_line(alpha = 0.5) +
  scale_colour_manual(breaks = c("occluded", "visible"), 
                      na.value = "white",
                      values = c("grey75", "orange")) +
  expand_limits(y = c(0, 70))

```

```{r}
ggplot(subset(second_2024_latest.tb, 
              time > ymd_hm("2024-05-05 10:38") & time < ymd_hm("2024-05-05 10:40")),
       aes(time, PAR_umol)) +
  geom_line() +
  geom_point(aes(colour = solar_disk)) +
  scale_colour_manual(breaks = c("occluded", "visible"), 
                      na.value = "white",
                      values = c("grey75", "orange")) +
  expand_limits(y = 0)

```

```{r}
ggplot(subset(second_2024_latest.tb, 
              time > ymd_hm("2024-05-05 10:42") & time < ymd_hm("2024-05-05 10:44")),
       aes(time, PAR_umol)) +
  geom_line() +
  geom_point(aes(colour = solar_disk)) +
  scale_colour_manual(breaks = c("occluded", "visible"), 
                      na.value = "white",
                      values = c("grey75", "orange")) +
  expand_limits(y = 0)

```

```{r}
ggplot(subset(second_2024_latest.tb, 
              time > ymd_hm("2024-05-05 14:11") & time < ymd_hm("2024-05-05 14:13")),
       aes(time, PAR_umol)) +
  geom_line() +
  geom_point(aes(colour = solar_disk)) +
  scale_colour_manual(breaks = c("occluded", "visible"), 
                      na.value = "white",
                      values = c("grey75", "orange")) +
  expand_limits(y = 0)

```

```{r}
ggplot(subset(second_2024_latest.tb, 
              time > ymd_hm("2024-04-30 10:42") & time < ymd_hm("2024-04-30 10:44")),
       aes(time, PAR_umol)) +
  geom_line() +
  geom_point(aes(colour = solar_disk)) +
  scale_colour_manual(breaks = c("occluded", "visible"), 
                      na.value = "white",
                      values = c("grey75", "orange")) +
  expand_limits(y = 0)

```
