---
title: "Diffuse fraction under clear sky"
subtitle: "Data simulated with the TUV model"
author: "Pedro J. Aphalo"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
rm(list = ls(pattern = "*"))
knitr::opts_chunk$set(tidy = FALSE, cache = FALSE)
```

When new data is available, first run 'read-data.Rmd', possibly after clearing the 'knitr' cache.

```{r}
library(dplyr)
library(ggplot2)
library(lubridate)
library(photobiology)
library(photobiologyInOut)
library(photobiologyWavebands)

theme_set(theme_bw())

options(warnPartialMatchAttr = FALSE,
        warnPartialMatchDollar = FALSE,
        warnPartialMatchArgs = FALSE)
```

## Diffuse PAR fraction from TUV

Read Quick TUV output files for different solar zenith angles and split the
components of total irradiance into separate spectra. We select files with
`low` in their names, with low levels of aerosols that seem match better the
data as can be seen in the next section. One can change `"low"` by `"4str"`
for the default aerosols from Quick TUV.

```{r, message = FALSE, warning = FALSE}

files <- list.files(pattern = "tuv", full.names = TRUE)

SZA.mspct <- source_mspct()

for (f in files) {
  name <- gsub("tuv-default-|-4str.htm|-low.htm", "", basename(f))
  print(name)
  SZA.mspct[[name]] <- read_qtuv_txt(f, ozone.du = 300)
}

SZA_tot.mspct <- source_mspct()
SZA_diff.mspct <- source_mspct()
SZA_dir.mspct <- source_mspct()
for (n in names(SZA.mspct)) {
  SZA_tot.mspct[[n]] <-
    SZA.mspct[[n]][ , c("w.length", "s.e.irrad")]
  SZA_diff.mspct[[n]] <-
    with(SZA.mspct[[n]], source_spct(w.length = w.length, s.e.irrad = s.e.irrad.diff.down))
  SZA_dir.mspct[[n]] <-
    with(SZA.mspct[[n]], source_spct(w.length = w.length, s.e.irrad = s.e.irrad.dir))
}
```

Compute PAR irradiances and the diffuse fraction for different SZAs.

```{r}
q_irrad(SZA.mspct, PAR(), scale.factor = 1e6)
PAR_tot.tb <- q_irrad(SZA_tot.mspct, PAR(), scale.factor = 1e6)
PAR_dir.tb <- q_irrad(SZA_dir.mspct, PAR(), scale.factor = 1e6)
PAR_diff.tb <- q_irrad(SZA_diff.mspct, PAR(), scale.factor = 1e6)

PAR.tb <-
  data.frame(
    SZA = as.numeric(gsub("tuv-|\\.htm", "", PAR_tot.tb$spct.idx)),
    SEA = 90 - as.numeric(gsub("tuv-|\\.htm", "", PAR_tot.tb$spct.idx)),
    total = PAR_tot.tb$`Q_PAR[`,
    direct = PAR_dir.tb$`Q_PAR[`,
    diffuse = PAR_diff.tb$`Q_PAR[`,
    diff_fraction = PAR_diff.tb$`Q_PAR[` / PAR_tot.tb$`Q_PAR[`
  )
```

Plot to check that data look good. We use solar elevation angle for the plot.

```{r}
ggplot(PAR.tb, aes(SEA, diff_fraction)) +
  geom_line() +
  expand_limits(y = 0)
```
```{r}
# We fit a spline obtaining a function that can be used to obtain by interpolation
# estimates of the diffuse PAR fraction for any solar elevation covered by the
# simulations.

clear_sky_PAR_diff_frac.fun <- splinefun(PAR.tb$SEA, PAR.tb$diff_fraction)
```

```{r}
ggplot(PAR.tb, aes(SEA, diff_fraction)) +
  stat_function(fun = clear_sky_PAR_diff_frac.fun, xlim = c(-10, 90), colour = "black") +
  geom_point(na.rm = TRUE, colour =  "orange") +
  scale_x_continuous(name = "Solar elevation angle (degrees)",
                     breaks = c(-10, 0, 15, 30, 45, 60, 75, 90)) +
  labs(y = expression("Diffuse fraction, "*Q[s] / Q[t])) +
  expand_limits(y = 0)

```
```{r}
saveRDS(clear_sky_PAR_diff_frac.fun, file = "spline-fun.RDS")
```

